---
title: "IODS Final assignment"
author: 
  name: "Elsa Isabel de Oliveira Marques" 
  email: "elsa.marques@helsinki.fi"
date: "10/12/2017"
abstract:
  An ‘abstract’ at the beginning of the page with a summary of your analysis (max 2 points)
  
output:
  html_document:    
    code_folding: hide
    fig_height: 4
    fig_width: 6
    highlight: espresso
    number_sections: yes
    toc: yes
    toc_float: yes 
---

***
  
```{r child = "Research topic.Rmd"}
```
  
***


 Also you can use these datasets from the Factominer-package to do Multiple Correspondence Analysis:

Hobbies


- MCA (on the categorical variables, perhaps categorize more variables?)

- Linear regression (for the grades variable, for example)

- LDA (on categorized grades for example)





# - Load data and explore data summary and dimensions 
```{r, echo = FALSE}

```

{r cols.print=3, rows.print=3}
mtcars


# - Data Analysis 

Visually clear and interesting explorations of the variables of interest in the data, from the point of view of your research question. Include interpretations of the distributions and relationships of the variables. Use captions to draw the reader’s focus on the interesting parts of your tables and graphics. (max 8 points)

Brief description of the method(s) you are using in your own words (max 3 points)

Presentation of the results of your analysis including visualizations and summaries and a thorough interpretation of the results including a validation analysis of the method. (max 16 points)

## - llk

## - adada


# - Conclusions and discussion 

Conclusions and discussion (max 2 points)



Visually clear and interesting explorations of the variables of interest in the data, from the point of view of your research question. Include interpretations of the distributions and relationships of the variables. Use captions to draw the reader’s focus on the interesting parts of your tables and graphics. (max 8 points)

Brief description of the method(s) you are using in your own words (max 3 points)

Presentation of the results of your analysis including visualizations and summaries and a thorough interpretation of the results including a validation analysis of the method. (max 16 points)




