---
title: "MCA"
author: "Elsa Marques"
date: "12/11/2017"
output: 
  html_document:
    code_folding: hide
    fig_height: 4
    fig_width: 6
    highlight: espresso
    number_sections: yes
    toc: yes
    toc_float: yes 
---

 > Visually clear and interesting explorations of the variables of interest in the data, from the point of view of your research question. Include interpretations of the distributions and relationships of the variables. Use captions to draw the reader’s focus on the interesting parts of your tables and graphics. (max 8 points)
 
 > Presentation of the results of your analysis including visualizations and summaries and a thorough interpretation of the results including a validation analysis of the method. (max 16 points)
 
 
# - Data Analysis 

## - Load data and explore data summary and dimensions 
```{r, echo = FALSE}
Hobbies <- read.table("hobbies_.csv", header = TRUE, Sep = ";")
```

{r cols.print=3, rows.print=3}
mtcars




```{r, echo = FALSE}
# visualize the dataset
gather(tea_time) %>% ggplot(aes(value)) + facet_wrap("key", scales = "free") + geom_bar() + theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6))
```



## - MCA


Brief description of the method(s) you are using in your own words (max 3 points)
```{r}
res.mca <- MCA(hobbies, quali.sup=19:22, quanti.sup=23, method="Burt")

plot(res.mca,invisible=c("ind","quali.sup"),hab="quali") ### active var. only
plot(res.mca,invisible=c("var","quali.sup"),cex=.5,label="none") ### individuals only
plot(res.mca,invisible=c("ind","var"),hab="quali") ### supp. qualitative var. only

dimdesc(res.mca)
plotellipses(res.mca,keepvar=1:4)
```


```{r, echo = FALSE}
# multiple correspondence analysis
mca <- MCA(hobbies_, graph = FALSE)

# summary of the model
summary(mca)
```

(on the categorical variables, perhaps categorize more variables?)




```{r, echo = FALSE}
# visualize MCA
plot(mca, invisible=c("ind"), habillage = "quali")
```

 - Now upon removal of the individuals in the plot you can better verify how the variables better relate with one or the other dimension. 
 
  - The typical MCA bipolot shows the original classes of the discrete variables on the same "map". This allows us to check the data distributions of the different variables, revealing the connections (correspondences) between different things that would be quite impossible to see from the corresponding cross table. The plot including all the data seems extremelly busy and difficult to verify the correlations between the variables 
  
  


```{r, echo = FALSE}
res.mca <- MCA(hobbies,quali.sup = 19:22, quanti.sup = 23)
```





